Data Cleaning

### 2025-12-09 Data Cleaning.Rmd
Completed preliminary classification algorithm analysis via random forest, LDA, QDA.

@@

class_df <- rwar_dosing_lag1

threshold_smallchange <- 0.01

@@

# # Temp
# lagged_data$Temp_lag1[1] <- lagged_data$Temp_lag1[2]
# lagged_data$Temp_lag2[1:2] <- lagged_data$Temp_lag2[3]
# # pH
# lagged_data$pH_lag1[1] <- lagged_data$pH_lag1[2]
# lagged_data$pH_lag2[1:2] <- lagged_data$pH_lag2[3]
# # Conductivity
# lagged_data$Cond_lag1[1] <- lagged_data$Cond_lag1[2]
# lagged_data$Cond_lag2[1:2] <- lagged_data$Cond_lag2[3]
# # Turbidity
# lagged_data$Turb_lag1[1] <- lagged_data$Turb_lag1[2]
# lagged_data$Turb_lag2[1:2] <- lagged_data$Turb_lag2[3]

lm_fit <- lm(Ferric.chloride..mg.L ~ ., data=lagged_data1 |> select(-c(Timestamp)))
summary(lm_fit)

@@

## TRAIN/TEST DATA
train_df <- df_lag_incdec |> filter(year(Timestamp) <= 2019)
test_df  <- df_lag_incdec |> filter(year(Timestamp) == 2020)

## RANDOM FOREST

rf_spec <- rand_forest(mtry = sqrt(.cols())) |>
  set_engine("randomForest", importance = TRUE) |>
  set_mode("classification")

rf_fit <- rf_spec |>
  fit(high_sales ~ ., data = df_train)

vip(rf_fit) ## still Price and Shelf location are the most important

rf_fit |>
  augment(new_data = df_test) |>
  accuracy(truth = high_sales, estimate = .pred_class)

# 5. Compare the OOB confusion matrix to your test confusion matrix. [**Hint:** The `confusion` element of the model output fit is OOB.]

## OOB:
rf_fit |>
  extract_fit_engine() %>%
  .$confusion

## test confusion
rf_fit |>
  augment(new_data = df_test) |>
  conf_mat(truth = high_sales, estimate = .pred_class)
```

>>>>

```{r, include=F}
library(ISLR) ## data package
library(tidyverse) ## data manipulation
library(tidymodels) ## tidy modeling
library(knitr) ## tables
library(rpart.plot) ## tree diagrams
library(vip) ## plotting variable importance
library(dplyr)
library(tidyr)
library(readr)
library(lubridate)
library(ggplot2)
library(zoo)
library(purrr)
library(knitr)
library(patchwork) ## plot arrangement
library(parsnip)
library(discrim)
```

@@

## Set up a column of inc/dec for 
class_df <- rwar_dosing_lag1
threshold_smallchange <- 0.01 # cutoff for dose changes that are approx. zero

@@

@@

```{r}
## SET SEED
set.seed(445)

## TRAIN/TEST DATA (2018-2019 and 2020)
df_full <- df_lag_incdec |> select(-c(FeCl_dose, Polymer_dose))
df_train <- df_full |> filter(year(Timestamp) <= 2019)
df_test  <- df_full |> filter(year(Timestamp) == 2020)
```

```{r}
set.seed(445)

## RANDOM FOREST
rf_spec <- rand_forest(mtry = sqrt(.cols())) |>
  set_engine("randomForest", importance = TRUE) |>
  set_mode("classification")

# Different model fits
rf_fit1 <- rf_spec |>
  fit(FeCl_change ~ ., data = df_train |> 
        select(-c(Polymer_change, Timestamp)))
rf_fit2 <- rf_spec |>
  fit(FeCl_change ~ mean_Temp + mean_pH + Temp_lag1 + Temp_lag2 + pH_lag1 + pH_lag2, 
      data = df_train)
rf_fit3 <- rf_spec |>
  fit(FeCl_change ~ mean_pH + mean_Temp, 
      data = df_train)

# Model fit to take performance metrics of:
rf_fit <- rf_fit3

# Most important predictors
vip(rf_fit)

## Accuracy
rf_fit |>
  augment(new_data = df_test) |>
  accuracy(truth = FeCl_change, estimate = .pred_class)
## Test confusion matrix
rf_fit |>
  augment(new_data = df_test) |>
  conf_mat(truth = FeCl_change, estimate = .pred_class)

## Notes
# 34% accuracy for all predictors
# 47% accuracy for just temperature and pH

```{r}
## set seed
set.seed(445)

## LDA
lda_spec <- discrim_linear()
lda_fit1 <- lda_spec |>
  fit(FeCl_change ~ ., data = df_train |> 
        select(-c(Polymer_change, Timestamp)))
lda_fit2 <- lda_spec |>
  fit(FeCl_change ~ mean_pH + mean_Temp, data = df_train)

# select LDA to take metrics of
lda_fit <- lda_fit2

# Test confusion matrix
lda_fit |> 
  augment(new_data = df_test) |>
  conf_mat(truth = FeCl_change, estimate = .pred_class)

# Test accuracy
lda_fit |>
  augment(new_data = df_test) |>
  accuracy(truth = FeCl_change, estimate = .pred_class)

## Notes
# 39% test accuracy on all predictors
# 41% test accuracy on temp and pH

## QDA
qda_spec <- discrim_quad()
qda_fit1 <- qda_spec |>
  fit(FeCl_change ~ ., data = df_train |> 
        select(-c(Polymer_change, Timestamp)))
qda_fit2 <- qda_spec |>
  fit(FeCl_change ~ mean_pH + mean_Temp, data = df_train)

# select LDA to take metrics of
qda_fit <- qda_fit2

# Test confusion matrix
qda_fit |> 
  augment(new_data = df_test) |>
  conf_mat(truth = FeCl_change, estimate = .pred_class)

# Test accuracy
qda_fit |>
  augment(new_data = df_test) |>
  accuracy(truth = FeCl_change, estimate = .pred_class)

## Notes
# 39.5% test accuracy on all predictors
# 43.7% test accuracy on temp and pH
